---
title: "Seafood Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
    #runtime: shiny
    output: html_document
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(lubridate)
library(vroom)
library(tidyverse)
#library(shiny)
library(DT)
library(crosstalk)
library(kableExtra)

```


```{r global, include=FALSE}
## Loading data
df_Seafood <- vroom("Seafood-database.csv")

## processing data
# Turning Character columns in Factors. But first selecting all columns of class = Character
# https://stackoverflow.com/questions/33180058/coerce-multiple-columns-to-factors-at-once
df_Seafood[map_lgl(df_Seafood, is.character)] <- lapply(df_Seafood[map_lgl(df_Seafood, is.character)], factor)

#### turning Year in to date.year
## Issue here. We only have years, and that is not a valid date format for plotting. But, since the values are annual sum. But, for plotting it looks better using the 1st of Jan. If we would use e.g. the 31.12.2019, it would look like the data is for 2020 :( 

df_Seafood$Year <-dmy(as.character(paste("01-01",df_Seafood$Year)))

#df_Seafood_shared <- SharedData$new(df_Seafood)

df_UKPOP <- vroom("ONS_UK_population-19712020.csv")
df_UKPOP$Year <-dmy(as.character(paste("01-01",df_UKPOP$CDID)))



```

```{r sunburst data wrangling, include=FALSE}
### We are trying to build a sunburst plot from our import export dataset
## the data needs to have the following structure ids, labels, parents, values
## values need to be the total of each ids e.g. Export all fish exported

## Start with import and exports

df_Seafood %>% group_by(ids = Commodity) %>%  summarise(values = sum(na.omit(Volume)))
df_Seafood %>% group_by(parents = Commodity, ids = SpeciesType) %>%  summarise(values = sum(na.omit(Volume)))
df_Seafood %>% group_by(Commodity, parents = SpeciesType, ids = Species) %>%  summarise(values = sum(na.omit(Volume)))


## Import & Export sums
df_sunburst <- df_Seafood %>% 
  group_by(ids = as.character(Commodity)) %>% 
  summarise(values = sum(na.omit(Volume))) %>% 
  mutate(labels = ids, parents ="")


## Species Type sums
df_sunburst.temp <-df_Seafood %>% 
  group_by(parents = as.character(Commodity), labels = as.character(SpeciesType)) %>% 
  summarise(values = sum(na.omit(Volume))) %>% 
  mutate(ids = paste(parents," - ", labels)) # to create unique ids


df_sunburst <- rbind(df_sunburst,df_sunburst.temp)

## Species sums
df_sunburst.temp <- df_Seafood %>% 
  group_by(Commodity, SpeciesType, labels = as.character(Species)) %>% 
  summarise(values = sum(na.omit(Volume))) %>% 
  ungroup() %>% 
  mutate(parents = paste(Commodity," - ", SpeciesType),
         ids = paste(Commodity," - ", SpeciesType, " - ", labels)) %>% 
  select(ids, labels, parents, values)

df_sunburst <- rbind(df_sunburst,df_sunburst.temp)

### last one years (no calucation)
df_sunburst.temp <- df_Seafood %>% 
  select(Commodity, SpeciesType, Species, Year, Volume) %>% 
  mutate(parents = paste(Commodity," - ", SpeciesType, " - ", Species),
         ids = paste(Commodity," - ", SpeciesType, " - ", Species, " - ", Year),
         labels = lubridate::year(Year),
         values = Volume) %>% 
  select(ids, labels, parents, values)

df_sunburst <- rbind(df_sunburst,df_sunburst.temp)

df_Seafood.cp <- df_Seafood %>% filter(Species %in% c("Cod", "Haddock", "Prawns", "Tuna", "Salmon", "Mackerel", "Herring"))

## "Consumption" & "Purchases sums
df_sunburst.cp <- df_Seafood.cp %>% 
  group_by(ids = as.character(Commodity)) %>% 
   summarise(values = sum(na.omit(Volume))) %>% 
  mutate(labels = ids, parents ="")


## Species Type sums
df_sunburst.temp <-df_Seafood.cp %>% 
  group_by(parents = as.character(Commodity), labels = as.character(SpeciesType)) %>% 
   summarise(values = sum(na.omit(Volume))) %>% 
  mutate(ids = paste(parents," - ", labels)) # to create unique ids


df_sunburst.cp <- rbind(df_sunburst.cp,df_sunburst.temp)

## Species sums
df_sunburst.temp <- df_Seafood.cp %>% 
  group_by(Commodity, SpeciesType, labels = as.character(Species)) %>% 
   summarise(values = sum(na.omit(Volume))) %>% 
  ungroup() %>% 
  mutate(parents = paste(Commodity," - ", SpeciesType),
         ids = paste(Commodity," - ", SpeciesType, " - ", labels)) %>% 
  select(ids, labels, parents, values)

df_sunburst.cp <- rbind(df_sunburst.cp,df_sunburst.temp)

### last one years (no calucation)
df_sunburst.temp <- df_Seafood.cp %>% 
  select(Commodity, SpeciesType, Species, Year, Volume) %>% 
  mutate(parents = paste(Commodity," - ", SpeciesType, " - ", Species),
         ids = paste(Commodity," - ", SpeciesType, " - ", Species, " - ", Year),
         labels = lubridate::year(Year),
         values = Volume) %>% 
  select(ids, labels, parents, values)

df_sunburst.cp <- rbind(df_sunburst.cp,df_sunburst.temp)


```

```{r table Data availablity}

df_SF_availablity <- df_Seafood %>% select(Species, Commodity) %>%
  unique() %>% 
  mutate(Available = "Yes") %>% 
  pivot_wider(names_from = Commodity, 
              values_from = Available,
              values_fill = "No") %>% 
  mutate(Exports = as.factor(Exports),
         Imports = as.factor(Imports),
         Production = as.factor(Production),
         Consumption = as.factor(Consumption),
         Purchases = as.factor(Purchases))



```

```{r all species through year}

df_figure <- df_Seafood %>% select(Commodity, Species, SACN, Year, Volume, Units) %>% group_by(Commodity, Species, SACN, Year) %>% summarise(Volume = sum(Volume)) %>% mutate(Year_notDate = as.numeric(year(Year))) %>% left_join(df_UKPOP %>%  select(CDID, UKPOP), by = join_by(Year_notDate == CDID)) %>% mutate( Volume = Volume/UKPOP) %>% pivot_wider(names_from = Commodity, values_from = Volume)

#Subplots for chart 3
fig1 <-df_figure %>%  plot_ly(type = 'bar', x=~Year, y=~Exports, color = ~Species, legendgroup =~Species) %>% 
  layout(yaxis = list(title = "Volume [g/Capita]"))
fig2 <-df_figure %>%  plot_ly(type = 'bar', x=~Year, y=~Imports, color = ~Species, legendgroup =~Species, showlegend= FALSE)%>% 
  layout(yaxis = list(title = "Volume [g/Capita]"))
fig3 <-df_figure %>%  plot_ly(type = 'bar', x=~Year, y=~Production, color = ~Species, legendgroup =~Species, showlegend= FALSE)%>% 
  layout(yaxis = list(title = "Volume [g/Capita]"))
fig4 <-df_figure %>%  plot_ly(type = 'bar', x=~Year, y=~Consumption, color = ~Species, legendgroup =~Species, showlegend= FALSE) %>% 
  layout(yaxis = list(title = "Volume [g/Capita]"))
fig5 <-df_figure %>%  plot_ly(type = 'bar', x=~Year, y=~Purchases, color = ~Species, legendgroup =~Species, showlegend= FALSE) %>% 
  layout(yaxis = list(title = "Volume [g/Capita]"), xaxis = list(title = "Years"))

```


Dashboard
=======================================================================

## Row {.sidebar}



## Row {.tabset .tabset-fade}
### Table 1 Data availablitiy

Table 1 shows which commodity data is available for which species in the seafood database.The Seafood data base contains a decade worth (2009 - 2020) of data for 88 marine species. When the data availability is __"No"__ it indicates that the data is either not captured, not available to us or is not relevant for the respective species, e.g. not every species can be farmed in aquaculture. 

```{r table 1 kable color }
df_SF_availablity %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "responsive")) %>% 
  column_spec(2, color = "white", background = ifelse(df_SF_availablity$Exports == "Yes",  "green", "red")) %>% 
  column_spec(3, color = "white", background = ifelse(df_SF_availablity$Imports == "Yes", "green", "red")) %>% 
  column_spec(4, color = "white", background = ifelse(df_SF_availablity$Production == "Yes", "green", "red")) %>% 
  column_spec(5, color = "white", background = ifelse(df_SF_availablity$Consumption == "Yes", "green", "red")) %>% 
  column_spec(6, color = "white", background = ifelse(df_SF_availablity$Purchases == "Yes", "green", "red"))



```


### Chart 1 Sunburst chart

```{r Chart 1 Sunburst chart}
layout(plot_ly(df_sunburst,
               ids = ~ids,
               labels = ~labels,
               parents = ~parents,
               values = ~values,
               type = 'sunburst',
               branchvalues = 'total'
), colorway = c("#009E73", "#0072B2", "#56B4E9" ,"#E69F00", "#F0E442" ))

```

### Chart 2 Sunburst 5 Species

```{r Chart 2 Sunburst Consumption & Purchase}

layout(plot_ly(df_sunburst,
               ids = ~ids,
               labels = ~labels,
               parents = ~parents,
               values = ~values,
               type = 'sunburst',
               branchvalues = 'total'
), colorway = c("#009E73", "#0072B2", "#56B4E9" ,"#E69F00", "#F0E442" ))
```

### Chart 3 Commodity data

```{r Chart 3 Commodity data}
#help for check out https://plotly.com/r/subplots/

subplot(fig1, fig2, fig3, fig4, fig5, nrows = 5, shareX = TRUE, titleY = TRUE) %>% 
  layout(yaxis = list(
    zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'),
    annotations = list( 
  list( 
    x = 0.1,  
    y = 1.0,  
    text = "UK Exports",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.1,  
    y = 0.79,  
    text = "UK Imports",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.1,  
    y = 0.58,  
    text = "UK Production",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.1,  
    y = 0.39,  
    text = "UK Consumption",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
   list( 
    x = 0.1,  
    y = 0.19,  
    text = "UK Purchases",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ) )
  )

 
  


```




```{r Chart 5 Sankey, eval=FALSE, include=FALSE}

df_test <- df_Seafood %>% select(Commodity, SACN, Year, Volume) %>% filter(Year == "2018-01-01") %>% 
  filter(Commodity %in% c("Exports","Imports", "Production", "Purchases")) %>%
  group_by(SACN, Commodity) %>% summarise(Volume = sum(Volume, na.rm = TRUE))



plot_ly(
  type = "sankey",
  orientation = "h",
  node = list(
    label = c("Product"	,"Import",	"Lean",	"Oily",	"Other",	"Shellfish",	"Export",	"Purchase"),
    color = c("green", "red", "blue", "blue", "blue", "blue","lightblue", "lightgreen"),
    pad = 15,
    thickness = 20,
    line = list(
      color = "black",
      width = 0.5
    )
  ),
  
  link = list(
    source = c(2,1,0,2,3,1,0,3,4,1,0,4,5,1,0,5),
    target = c(6,2,2,7,6,3,3,7,6,4,4,7,6,5,5,7),
    value =  df_test$Volume
  )
)


```

Key Messages
=======================================================================

```{r Key Messages}
tbl_key_num <- c(1,2,3,4)
tbl_key_text <- c("This unique database harmonises secondary datasets across the whole seafood supply chain, to species level, between 2009 and 2020 (the most up-to-date data available) for the United Kingdom (UK).",
                 "Datasets include seafood production (capture and aquaculture), trade (imports and exports), purchases (within and out-of-home), and human consumption.",
                 "Data are available annually (calendar year), for up to 73 aquatic species. Species can be aggregated according to the Food and Agricultural Organization (FAO) 'International Standard Statistical Classification for Aquatic Animals and Plants' (ISSCAAP) and the Scientific Advisory Committee in Nutrition (SACN) classification.",
                 "Corresponding nutrient values (omega-3 fatty acids, vitamin A, vitamin B12, vitamin D, calcium, iodine, iron, zinc, and selenium) for all aquatic species are provided, where possible, to allow mapping of the dietary flow of nutrients.")
tbl_key_blank <- c("","","","")

tbl_key_msg <- data.frame(tbl_key_num,tbl_key_blank,tbl_key_text,tbl_key_blank)


tbl_key_msg %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(c("striped","hover"), full_width = FALSE,
              html_font = "Cambria",
              font_size = 18)
```




About
=======================================================================


```{r About}
tbl_about <-c("This dashboard aims to understand how our consumption of seafood relates to food supply chains - what is being produced, what is imported and exported, and how does this relate to what we eat and what we should be eating for optimal health. This knowledge will help to reconsider national food supply chains, but also ensure that current national dietary guidelines take into account consumer health outcomes, as well as food supplies and the sustainability of food production systems.",
              "",
              "Created by [Dr. Anneli Lofstedt](https://orcid.org/0000-0002-1865-0891) and [Dr. Bernhard Scheliga](https://orcid.org/0000-0003-2764-6605).",
              "",
      "Funded by the Rural and Environment Science and Analytical Services Division (RESAS) of the Scottish Government, project RI-B5-04."
              )

tbl_about %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(c("striped","hover"), full_width = FALSE,
              html_font = "Cambria",
              font_size = 18)


```


Source
=======================================================================



```{r Source table version 2}
### Version 2
tbl_SourceName <- c("Capture:", "Aquaculture:","Imports and Exports:", "Within and out-of-home:","National Diet and Nutrition Survey", "Nutrient Analysis of Fish","Food Composition Table")
tbl_URL <- c("[UK sea fisheries annual statistics - GOV.UK (www.gov.uk)](https://www.gov.uk/government/collections/uk-sea-fisheries-annual-statistics)", 
             "[Database - Eurostat (europa.eu)](https://ec.europa.eu/eurostat/web/main/data/database)",
             "[UK Trade Info - GOV.UK (www.gov.uk)](https://www.uktradeinfo.com/trade-data/)",
             "[Family food statistics - GOV.UK (www.gov.uk)](https://www.gov.uk/government/collections/family-food-statistics)",
             "[National Diet and Nutrition Survey - GOV.UK (www.gov.uk)](https://www.gov.uk/government/collections/national-diet-and-nutrition-survey)",
             "[Nutrient Analysis of Fish - GOV.UK (www.gov.uk)](www.gov.uk)",
             "[Norwegian Food Composition Table (matvaretabellen.no)](https://www.matvaretabellen.no/about-us/)")
tbl_Source <- data.frame(tbl_SourceName,tbl_URL)

### adding footnote to Aquaculture
tbl_Source[2,1]  <- paste0(tbl_Source[2,1], 
       footnote_marker_symbol(1))


tbl_Source %>% kbl(col.names = NULL, escape = FALSE) %>% 
  kable_paper("hover", full_width = FALSE,
              position = "left",
               html_font = "Cambria",
              font_size = 18) %>% 
  pack_rows("Production data",1,2) %>% 
  pack_rows("Trade data",3,3) %>%
  pack_rows("Purchase data",4,4) %>%
  pack_rows("Consumption data",5,5) %>%
  pack_rows("Nutrient composition data",6,7) %>%
  footnote(symbol = c("Currently, data is available for download until 2018 (01/01/2024). More recent data was aquired via personal communications with CEFAS"))

```


Related Publication
=======================================================================

```{r Related Publication}
tbl_Rel_Pub <- c("[place holder]",
                "[PowerPoint Presentation (sefari.scot)](https://sefari.scot/sites/default/files/documents/RESAS_DynamicEarthPosterFinal.pdf) 2023 May")


tbl_Rel_Pub %>% kbl(col.names = NULL, escape = FALSE) %>% 
  kable_paper("hover", full_width = FALSE,
              position = "left",
               html_font = "Cambria",
              font_size = 18) %>% 
  pack_rows("Paper",1,1) %>% 
  pack_rows("Presentation",2,2)



```


Contact
=======================================================================


```{r}
tbl_contact <- c("[Prof. Baukje de Roos](https://www.abdn.ac.uk/people/b.deroos#about)",
                 "University of Aberdeen - [The Rowett Institute](https://www.abdn.ac.uk/rowett/)",
                 "[b.deroos@abdn.ac.uk](b.deroos@abdn.ac.uk)")
tbl_contact %>% kbl(col.names = NULL, escape = FALSE) %>% 
  kable_paper("hover", full_width = FALSE,
              position = "left",
               html_font = "Cambria",
              font_size = 18) %>% 
  pack_rows("Contact Information",1,3) 

```
